%!TEX root = Manuscript.tex

\silentchapter{Conclusion}
\label{chap:concl}
%\addcontentsline{toc}{chapter}{Conclusion}

In this thesis, we presented the problem of minimizing latency of periodic flows in a packet switched network. Current networks in use for telecommunication are based on statistical multiplexing: the links are dimensioned considering the average bit rate of the flows so that the flows can share a link most of the time or be buffered until enough capacity is available. Statistical multiplexing is a low-cost solution to deploy a network, but it does not guarantee the latency of the packets using it. If a burst of data is sent by one flow, shared resources become critical and some packets are buffered while waiting for their availability. These buffers are called contention buffers, and are a major source of latency.

We study the Cloud-RAN application case. In C-RAN, radio antennas periodically send packets to datacenters, that compute an answer and send it back to the antennas. The packets must have an end-to-end latency lower than a maximal value, required by the protocols. Statistical multiplexing is not able to guarantee an end-to-end latency for packets, and because of contention buffers, the more loaded is a network, the largest is the latency. In our C-RAN use case, the flows are periodic and a large amount of data is sent at each period by the antennas and the datacenters. Thus, managing the packets with statistical multiplexing is not appropriate.

Several working groups (DetNet, TSN, see Chapter~\ref{chap:TSN}), have developed standards and mechanisms ensuring an upper bound on the latency in packet switched networks. The network devices can reserve a port during a given time to forward the traffic of a given flow without contention buffer. The arrival date of the packets in a device of the network must then be known and precise. To do so, a scheduling of every output port of the devices is computed ahead. Current approaches to compute this scheduling are based on stochastic laws, since most of the internet traffic follows a stochastic behavior. In such a situation, it is impossible to completely get rid of contention buffers. Nevertheless, since our flows are deterministic (the amount of data and the periodicity of the packets remain the same all over time), we show that deterministic scheduling guarantees a minimal latency of deterministic flows and it also helps to reduce the latency of stochastic best-effort flows. 
Furthermore, TSN and DetNet mechanisms induce several sources of additional latency, like guard time around packets to prevent the time shift between clocks or buffering time of the header of the packets in every switch to read the destination. In this thesis, we go further by proposing solutions to reduce the end-to-end latency of packets to its physical transmission time by proposing a new generation of switch that get rid of technical constraints imposed by a statistical vision of the network.


 This thesis focuses on the problem of computing a deterministic scheduling for periodic flows, a problem that we prove to be $\NP$-hard for arbitrary networks.
 We study in Chapters~\ref{chap:PAZL}~and~\ref{chap:PALL} this problem when the flows are unsynchronized, that is, we can choose the emission date of the packets in the sources of the flows. This case does not perfectly match with Cloud-RAN, but it corresponds to various use cases, like Industry 4.0, autonomous vehicle\dots
In Chapter~\ref{chap:PAZL}, we give several greedy algorithms and one FPT algorithm that allows to reduce the latency to the physical transmission time (i.e. without any contention) on a common network topology with a single shared link, when the load induced by the deterministic traffic is low enough. We experimentally show using the exact FPT algorithm, that on very loaded networks (when the load is greater than $80\%$), it is not possible to get rid of contention buffers. 
We then propose solutions that buffer packets in nodes of the network, but we try to minimize this additional latency. Remark that in such an approach, the buffers are not anymore a consequence of contention that we cannot control or predict, but a tool to organize the packets.
In Chapter~\ref{chap:PALL}, we study the problem of organizing flows in a network with a single shared link (as in Chapter~\ref{chap:PAZL}), and allowing one buffer (positioned in the datacenters) on the route for every packet. We propose several greedy algorithms and one FPT algorithm based on a classical scheduling algorithm that we adapted for periodicity. The performances of our algorithm are excellent, we show it is possible to reduce latency to the physical transmission time of the longest route in $99,9\%$ of the cases, while statistical multiplexing, even prioritizing critical flows adds a latency to the flows, due to contention buffers, equal to $1/4$ of the period.

We study in Chapter~\ref{chap:SPALL} the C-RAN use case, in which all antennas send their messages at the same date, on arbitrary networks. We propose a compact form of the solutions to our problem, that allows to define a neighborhood of a solution. Then, several local search heuristics are designed using this notion of neighborhood. A branch and bound algorithm based 
on the compact form of the solutions is also proposed and run efficiently for small C-RAN network with ten to twenty routes. Then, we experimentally show that our approach dramatically over performs statistical multiplexing in terms of latency.

We then show in Chapter~\ref{chap:BE} how to adapt our algorithm not to impact best-effort flows latency while scheduling the C-RAN traffic. We explain how to adapt all our algorithms, by solving instances with artificially increased size of messages, in a way which does not impact the latency of C-RAN datagrams, and smooth the load of C-RAN traffic all over the period. 
 We show that, even if our approach induces an additional use of bandwidth due to resource reservation, we are able to improve the average latency of best-effort traffic while minimizing the C-RAN traffic latency. We also show similar results in an industrial optical ring, in which scheduling the C-RAN packets is trivial because of the multiplexing of the electronic signals over the optical ring.

We have proposed solutions to minimize the end-to-end latency in various use cases: Cloud-RAN, Industry 4.0, motion control, autonomous vehicle, etc\ldots
Reducing the transmission latency allows to:
\begin{itemize}
	\item Respect latency constraints required by protocols
	\item Increase the Network Quality of Service 
	\item Give more time to the other components of the chain (computation in datacenters for C-RAN example) 
	\item Augment the maximum physical links length, which means, for C-RAN; a wider area of development and thus lower exploitation and development costs (CAPEX, OPEX).
\end{itemize} 

\section*{Limits and Further researchs}

Several questions remain open in our work. We conjecture that \pazl and \pall are $\NP$-hard on star routed networks, but are not yet able to prove it. 
The algorithms we developed for \pazl and \pall are designed for star routed networks and most of them are not easy to extend to general networks. The greedy algorithms we proposed to solve \pazl can be adapted for arbitrary networks respecting the coherent routing property, but they can be proved to work only for small loads. These algorithms are not usable in their current state, and we need to study them carefully. Furthermore, the FPT algorithm we give to solve \minstra can be adapted to solve \pazl and \pall on arbitrary networks.
An experimental study has to be done to see if the approach is promising and the algorithm could be optimized for the case of unsynchronised messages.

This thesis arises in the context of SDN which aims to develop dynamical and programmable networks. In C-RAN for example, the radio network aims to be able to turn off antennas when the number of connected devices is low. Our algorithms for \pazl and \pall compute the scheduling for all flows, and must reschedule the entire solution if a flow is removed or added to the network to ensure a minimal latency. In the case of the algorithms presented for \minstra, the compact form of the solutions we presented allows us to efficiently add a flow to the best solution, but not to quickly re-compute the best solution when a flow is removed. A challenge is thus to design a dynamic algorithm, which can produce a new solution quickly after 
a local change in the network.

The measure we want to minimize is the maximal end-to-end latency of all flows. This often means that the flow using the longest route is never delayed in contention buffer, but the other flows are more or less impacted by the solutions we give. One can imagine use-cases in which the constraint on latency is not as strict, and where it makes more sense
to minimize the average latency of the flows. For this variant of the problem, the study of \pazl is still relevant, but it may be simpler from a complexiy point of view
and the algorithms may be quite different from the ones presented in this thesis.
 We can also expand our model to allow flows in the networks with different periods or sizes of message. Already several methods fail for mixing different message sizes, 
 since solving the problem \wta becomes \NP-hard in this context and mixing several periods require to completely change our algorithms.

In Section~\ref{sec:NGREEN} we introduce the fact that links of the networks may have different capacity. This induces additional latency due to the physical conversion. It could allow use cases requiring low latency to be developed on metropolitan networks. It is possible to adapt our model to take into account this issue, and to consider this physical conversion time while computing the solutions. 

Last, we consider that the routing is given in our model. It could be interesting to compute the routing 
along with the assignment to further improve the quality of the results. We plan to study this question on star routed networks, where the number of 
routings is limited and an exhaustive approach is conceivable. 
